Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
move figures around in alg section

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-0.7cm}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}[b]
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
      \code{($\lambda$ (a b) (foldr a b }\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b ($\lambda$ (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
  \label{fragmentExample}
\end{figure*}

>>>>

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}

\begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.

We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

\begin{figure}[b]
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
      \code{($\lambda$ (a b) (foldr a b }\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b ($\lambda$ (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
  \label{fragmentExample}
\end{figure}

To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
